Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2018-12-22 Manuscript/Figures/figure_legends.Rmd
WIP and issues on the current results

# Figure 4: Class Imbalance

Barplots illustrating the class imbalance problem featured across each of the four different resolutions that were analyzed. Minority classes represent the number of genomic bins that contain a TAD boundary, while the majority classes represent the number of genomic bins that do not contain a TAD boundary.

@@

# Figure 9: Comparing Performances

Comparing performances of random forest classification algorithms, across TAD boundary data at each resolution for models using distance-type predictors, with elastic-net regularization and random under-sampling. Performances were aggregated by taking the average of each metric across 50 iterations of random under-sampling.

# Figure 10: Comparing Variable Importance

Comparing variable importance for the top 15 most predictive functional genomic elements for each of the 4 different resolutions. The x-axis represents the standardized difference between the out-of-bag prediction accuracy after permuting each predictor variable, averaged across all trees. The greater the mean standardized difference, the more importance the predictor is to the model.

>>>>

**Figure 4: Techniques for addressing class imbalance.** Barplots illustrating the class imbalance problem featured across each of the four different resolutions that were analyzed. Minority classes represent the number of genomic bins that contain a TAD boundary, while the majority classes represent the number of genomic bins that do not contain a TAD boundary.

@@

**Figure 9: Comparing Performances** Comparing performances of random forest classification algorithms, across TAD boundary data at each resolution for models using distance-type predictors, with elastic-net regularization and random under-sampling. Performances were aggregated by taking the average of each metric across 50 iterations of random under-sampling. [??? Poor legend]

**Figure 10: Comparing Variable Importance.** Comparing variable importance for the top 15 most predictive functional genomic elements for each of the 4 different resolutions. The x-axis represents the standardized difference between the out-of-bag prediction accuracy after permutting each predictor variable, averaged across all trees. The greater the mean standardized difference, the more importance the predictor is to the model. [??? What are A/B/C/D panels?]

### 2018-12-22 Manuscript/manuscript.Rmd
WIP and issues on the current results

We first compared models with either overlap counts or overlap percent predictor types, from the 5kb resolution data (Figure 12). These results were from random forest models fit to data without any re-sampling, using either LASSO or Elastic-Net regulariations. We found that differences in performances across variable reduction techniques were marginal. More importantly, when evaluating model performance using accuracy or AUC, the model appeared to perform relatively well. At second glance, however, when considering both the F1-score and MCC, we were able to assess just how poorly the model was performing (Supplementary Table 9).

@@

When evaluatating the 5kb resolution data for a randomly under-sampled training set, it was found that models obtained better classification performance for F1 and MCC metrics (Figure 14). The increase in performance was present across both overlap predictor types. Models also featured a decrease in both accuracy and AUC, which was attributed to less inflated false negative rates (Supplementary Table 11). Variable reduction was found not to have an affect on resampling, and subsequent performances, as any differences were again marginal. Performances were also improved among the models using the other data resolutions as well. However, we found that model improvement dissipated as resolution increased (Supplementary Figure 15; Supplementary Table 12).

@@

@@

Finally, we directly compared the performances of models across the different resolutions. The comparisons consisted of random forest models using distance type predictors, elastic-net regularization, and 50 iterations of random under-sampling. The model using 5kb resolution domain data was found to perform slightly better compared to the 25kb and 50kb resolutions (Figure 9). A sharp decrease in performance was seen in the 100kb resolution.

The differences in performances was most evident when comparing the variable importances across the 4 resolutions. We found that for the 5kb resolution, the most predictive elements included insulator proteins, DNase hypersensitive sites, the transcription factor CTCF, as well as histone modifications H3k9ac, H2az, and H3k4me2 (Figure 10 A). Although, the insulator protein is featured as a top predictor in 2 of the other 3 resolutions, the rest of the predictors were less consistent among each other and less in agreement with the literature (Figure 10 B-D). The results from Figures 9 and 10 confirm that the model results appear to degrade as resolution increases with respect to both prediction and the identification of genomic elements associated with TAD boundaries.

>>>>

We first compared models with either overlap counts or overlap percent predictor types [??? Why not distance?], from the 5kb resolution data (Figure 12). These results were from random forest models fit to data without any re-sampling, using either LASSO or Elastic-Net regulariations. We found that differences in performances across variable reduction techniques were marginal. More importantly, when evaluating model performance using accuracy or AUC, the model appeared to perform relatively well. At second glance, however, when considering both the F1-score and MCC, we were able to assess just how poorly [??? Describe, why is it so?] the model was performing (Supplementary Table 9).

@@

Random under-sampling substantially improved classification performance as quantified by F1 and MCC metrics (Figure 14). The increase in performance was present across both overlap count and overlap percent predictor types [??? Again, where is distance?]. Models also featured a decrease in both accuracy and AUC, which was attributed to less inflated false negative rates (Supplementary Table 11). Variable reduction was found not to have an affect on resampling, and subsequent performances, as any differences were again marginal [??? Where is the proof?]. Notably, random under-sampling improved performance when using data at the other resolutions; however, at lower resolutions the performances were lower (Supplementary Figure 15; Supplementary Table 12). These results demonstrate that random under-sampling effectively alleviates class imbalance problem, especially for TAD boundaries detected from high-resolution Hi-C data.

@@

[??? Why not using all predictors together? Combine it together with the previous steps, testing no sampling, RUS, and the effect of distance]

@@

Finally, we directly compared the performances of models across the different resolutions. The comparisons consisted of random forest models using distance type predictors, elastic-net regularization, and 50 iterations of random under-sampling. The model using 5kb resolution domain data was found to perform slightly better compared to the 25kb and 50kb resolutions (Figure 9). A sharp decrease in performance was seen in the 100kb resolution. [??? This message is already evident from previous paragraphs and results. Not necessary to repeat it]

The differences in performances was most evident when comparing the variable importances across the 4 resolutions. [??? You never said and justified that you use distance predictor.] We found that for the 5kb resolution, the most predictive elements included insulator proteins, DNase hypersensitive sites, the transcription factor CTCF, as well as histone modifications H3k9ac, H2az, and H3k4me2 (Figure 10 A). Although, the insulator protein is featured as a top predictor in 2 of the other 3 resolutions, the rest of the predictors were less consistent among each other and less in agreement with the literature (Figure 10 B-D). The results from Figures 9 and 10 confirm that the model results appear to degrade as resolution increases with respect to both prediction and the identification of genomic elements associated with TAD boundaries.
